---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data


Let's create a dataframe called activitydata. 

We'll also create a true date variable (which contains both date and time).

```{r echo = TRUE}

activitydata <- read.table(file = unzip("activity.zip"),
                           header = TRUE,sep = ",")

activitydata$time <- as.character(activitydata$interval)


for(i in 1:nrow(activitydata)) {
  if (nchar(activitydata[i,"time"]) == 1) 
      {activitydata[i,"time"] = paste("000",activitydata[i,"time"],sep = "")
  } else if (nchar(activitydata[i,"time"]) == 2) 
      {activitydata[i,"time"] = paste("00",activitydata[i,"time"],sep = "")
  } else if (nchar(activitydata[i,"time"]) == 3) 
      {activitydata[i,"time"] = paste("0",activitydata[i,"time"],sep = "")
  } else 
      {activitydata[i,"time"] = activitydata[i,"time"]
  }
}

activitydata$datetime <- strptime(paste(activitydata$date,activitydata$time),
                                  "%Y-%m-%d %H%M")

```

## What is mean total number of steps taken per day?

Let's first aggregate data by date


```{r echo = TRUE}
activityaggregate <- aggregate(steps ~ date, data=activitydata,FUN =sum)
```

Let's then plot a histogram.

```{r histogram, fig.height=4}
hist(activityaggregate$steps, breaks = 10,  
     main = "Total steps taken by day", 
     xlab = "Steps")

```


And finally calculate the mean and median steps

```{r echo = TRUE}
mean(activityaggregate$steps)
median(activityaggregate$steps)
```


## What is the average daily activity pattern?

Let's plot the time series. 


```{r echo = TRUE}
plot(x=activitydata$datetime,y=activitydata$steps, type = "l",  xlab ="", 
     ylab = "Steps") 
```


The maximum number of steps happens in interval


```{r echo = TRUE}
activitydata[activitydata$steps == max(activitydata$steps,na.rm = TRUE) & is.na(activitydata$steps) == FALSE,c("datetime","steps")]
```


## Imputing missing values

Number of missing values
```{r echo = TRUE}
NAs <- is.na(activitydata$steps) 
sum(NAs)
```


We will use average steps by interval as the substitute value


```{r echo = TRUE}
avg_steps_per_interval <- mean(activitydata$steps,na.rm = TRUE)
```


Let's create new dataset and do the replacement of missing values


```{r echo = TRUE}
activitymodified <- activitydata

for(i in 1:nrow(activitymodified)) {
  if (is.na(activitymodified[i,"steps"])==TRUE) 
  {activitymodified[i,"steps"] = avg_steps_per_interval
  }
}

```



Let's aggregate the dataset with missing values replaced
and do the plotting.


```{r echo = TRUE}

activityaggregatemod <- aggregate(steps ~ date, data=activitymodified,FUN =sum)

hist(activityaggregatemod$steps,breaks = 10,  
     main = "Total steps taken by day (w/missing imputed)", 
     xlab = "Steps")


```


Let's calculate the median and mean.  There's actually virtually
no difference to the previous values.

```{r echo = TRUE}
mean(activityaggregatemod$steps)
median(activityaggregatemod$steps)

```

But there's naturally difference between the total steps taken.

```{r echo = TRUE}
sum((activityaggregatemod$steps))
sum((activityaggregate$steps))

```





## Are there differences in activity patterns between weekdays and weekends?

